A finite-state morphological transducer for Kyrgyz

### 2012-03-21 paper/paper.tex
precision and recall section

The results for precision and recall are presented in Table~\ref{table:precrecall}. That the precision is so high is not surprising: the transducer is rule-based and the test set is the corrected output of the analyser. In order to get a more accurate figure for precision, it would be necessary to have several linguists review the set. Low recall is due to forms which do not receive any analyses, in all cases this was due to missing stems in the lexicon. Out of the 1,000 words, 91 did not receive any analyses. Of these, 52 forms were proper names 28 were nouns, 6 were adjectives and the remaining 5 verbs. 
This would suggest that in terms of adding new stems, it would be a good idea to look at increasing both the number of proper names, and nouns.

\begin{table}[htbp]
	\centering
	\caption{Precision and recall of the analyser over the 1,000 word test set.}
	\begin{tabular}{rr}
		\toprule
		Precision & Recall \\
		\midrule 
		100 & 95.92  \\
		\bottomrule
	\end{tabular}
	\label{table:precrecall}
\end{table}

%FIXME: new stuff
The evaluation has brought to light some systematic issues with the transducer.  One such issue is that mono-syllabic verbs ending in vowels
(e.g., [же] \eng{eat}, [жуу] \eng{wash}, [оо] \eng{tilt})
do not take a gerund/infinitive in -/Uː/, but are only used with the morphologically, semantically, and syntactically similar -/(I)ш/ suffix.  To prevent these forms from being output, a different version of the current verbal continuation lexica should be made to capture the morphological difference---i.e., forms of these verbs that are marked as \texttt{<ger>} should be output with the -/(I)ш/ suffix instead of the -/Uː/ suffix.

>>>>

The results for precision and recall are presented in Table~\ref{table:precrecall}.

\begin{table}[htbp]
	\centering
	\caption{Precision and recall of the analyser over the 1,000 word test set.}
	\begin{tabular}{ll}
		\toprule
		Precision & Recall \\
		\midrule 
		97.32\% & 94.56\%  \\
		\bottomrule
	\end{tabular}
	\label{table:precrecall}
\end{table}

%FIXME: new stuff
That the precision is so high is not surprising: the transducer is rule-based and the test set is the corrected output of the analyser.  In order to get a more accurate figure for precision, it would be necessary to have several linguists review the set.
Our evaluation of the transducer brought to light some systematic ways that we could improve the transducer.

Low recall is due to forms which do not receive any analyses, in all cases this was due to missing stems in the lexicon. Out of the 1,000 words, 91 did not receive any analyses. Of these, 52 forms were proper names 28 were nouns, 6 were adjectives and the remaining 5 verbs.  This would suggest that in terms of adding new stems, it would be a good idea to look at increasing both the number of proper names, and nouns.

Low precision was due largely to incorrect stem categorisation.  Proofing our stem categorisation would increase precision significantly.  Besides this, several issues with the transducer were brought to light.  One such issue is that mono-syllabic verbs ending in vowels
(e.g., [же] \eng{eat}, [жуу] \eng{wash}, [оо] \eng{tilt})
do not take a gerund/infinitive in -/Uː/, but are only used with the morphologically, semantically, and syntactically similar -/(I)ш/ suffix.  To prevent these forms from being output, a different version of the current verbal continuation lexica should be made to capture the morphological difference---i.e., forms of these verbs that are marked as \texttt{<ger>} should be output with the -/(I)ш/ suffix instead of the -/Uː/ suffix.
